---
title: "PFU counting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installation 

Install packages if not installed yet

```{r eval = F}
# Install package. Enter "a" in console when prompted
if (!require("BiocManager", character.only = TRUE)) install.packages("BiocManager")
if (!require("EBImage", character.only = TRUE)) BiocManager::install("EBImage")
if (!require("tidyverse", character.only = TRUE)) install.packages("tidyverse")
if (!require("broom", character.only = TRUE)) install.packages("broom")
```

Load packages

```{r}
library(tidyverse)
library(broom)
library(EBImage)
```

# Note

- Be careful that this Rmd and the accompanied Rscript only work for the cropped image (see below) that only show ONE community at ONE dilution factor.

- This code only works for counting the total number of colony. I am not able to distinguish the different colony morphology yet. But I think for those with >100 colony this should help speed up a bit.

# Image processing

Read a test image. I have cropped the image such that each input image is a community at a dilution factor.

```{r}
# Read an image
img <- readImage(here::here("data/raw/test2.jpeg"))
display(img, method = "raster", all = T)
```

Reduce the dimension of the image from 3 to 1. Using the most contrast dimension

```{r}
# Grey scale
colorMode(img) = Grayscale
display(img, method = "raster", all = T)
```


```{r}
img <- img[,,3]
display(img, method = "raster")
```

Blur the image as the pre-treatment for gating out the background noise

```{r}
# Blur
img_bl <- gblur(img, sigma = 1)
display(img_bl, method = "raster")
```

Set the global threshold

```{r}
# Global threshold
threshold <- otsu(img_bl)
img_th <- combine(mapply(function(frame, th) frame > th, getFrames(img), threshold, SIMPLIFY=FALSE))
display(img_th, method = "raster")
```

Label the objects. Each object is a group of adjacent pixels.

```{r}
img_label = bwlabel(img_th)
```

Gating out the background. Be cautious that here I am setting a manual threshold (e.g., objects number of pixel <= 10 are considered noise). 

```{r}
# Remove background
temp <- img_label %>% table()
index_rm <- names(temp)[which(temp <= 10)] # Pixel size for gating out background
img_clean <- rmObjects(img_label, index_rm)
display(img_clean, method = "raster")
```

Watershed separates the adjacent colonies

```{r}
# Watershed
img_ws <- watershed(distmap(img_clean), 1)
display(colorLabels(img_ws), method = "raster")
```

You can get shape data for each colony: area, perimeter, radius, etc. The number of objects are the number of colonies.

```{r}
colony_shape <- computeFeatures.shape(img_ws) %>% as_tibble()
```

Number of total colonies

```{r}
nrow(colony_shape)
```

```{r}
colony_shape
```


# Test on another image

```{r}
# Read an image
img <- readImage(here::here("data/raw/test3.jpeg"))
display(img, method = "raster", all = T)

# Grey scale
colorMode(img) = Grayscale
display(img, method = "raster", all = T)
img <- img[,,3]
display(img, method = "raster")

# Blur
img_bl <- gblur(img, sigma = 1)
display(img_bl, method = "raster")

# Global threshold
threshold <- otsu(img_bl)
img_th <- combine(mapply(function(frame, th) frame > th, getFrames(img), threshold, SIMPLIFY=FALSE))
display(img_th, method = "raster")
img_label = bwlabel(img_th)
display(img_label, method = "raster")

# Remove background
temp <- img_label %>% table()
index_rm <- names(temp)[which(temp <= 10)] # Pixel size for gating out background
img_clean <- rmObjects(img_label, index_rm)
display(img_clean, method = "raster")

# Watershed
img_ws <- watershed(distmap(img_clean), 1)
display(colorLabels(img_ws), method = "raster")
```


In this case, the difference in colony size is very large. It makes sense to tell them apart by the size

```{r}
colony_shape <- computeFeatures.shape(img_ws) %>% as_tibble()
colony_shape
```

Relative count by clustering
```{r}
colony_shape %>%
    as.matrix() %>%
    kmeans(2) %>% 
    broom::tidy() %>%
    pull(size)
```

















